---
title: "TS_Project"
author: "Sherry Rodas"
date: "11/20/2018"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
suppressMessages(library(astsa))
suppressMessages(library(TSA))
suppressMessages(library(forecast))
suppressMessages(library(xts))
suppressMessages(library(tseries))
suppressMessages(library(dplyr))
```

Introduction

For the Time Series project we plan on using the Home Listings and Sales data provided by Zillow to generate a model that can predict monthly housing prices in a particular zip code. The data provides the median sale price for various housing types by zip code, starting from April 1996 all the way through September 2018.

We wanted to see how much home sales have changed close to the university campus, so we chose to narrow down the data to only include listings in the 77019 zip code. We performed EDA on the 1 Bedroom, 2 Bedroom, and 3 Bedroom housing types, but decided to use the data for the 1 Bedroom listings to build our model since it had the least variance out of the 3 datasets.

From the data provided by Zillow, we have built a time series object that ranges from April 1996 to May 2018. The model we build should predict the next 12 months, 4 months of which we will be able to compare actual prices to those predicted by our model (June-September 2018).

Besides being able to build a model that can predict listing sales prices, we would like to gain a better understanding of the trend, signals, and seasonality of the data. From prior apartment renting knowledge, we hypothesize the data will be seasonal and that the prices will go up in the summer months.


1 Bedroom House Prices
```{r}
bed1 = read.csv("~/Downloads/1bedroom_tp.csv", sep = ",", header = TRUE)
colnames(bed1) = gsub("[\\X,]","",colnames(bed1))
colnames(bed1)[1] = "date"
bed1 = bed1[,colSums(is.na(bed1)) == 0]
bed1_var = sapply(bed1[-1], var)
bed1_var_l = head(sort(bed1_var),20)
```

Choosing 77019 Zipcode
```{r}
b1_70019 = bed1 %>% select(date, `77019`)
b1_compare = b1_70019[267:270,]
b1_70019 = b1_70019[1:266,]
b1_70019$date = as.yearmon(b1_70019$date)
b1_ts = ts(b1_70019$`77019`, frequency = 12, start = c(1,4))
```


Results and Discussion

The following is the work we performed to accomplish our objective and the results we found through extensive analysis of the 1 Bedroom Time Series.

Plotting the Time Series Object allows a first look at trend or seasonality within the data.
```{r}
ts.plot(b1_ts, main="1 Bedroom Listing Sales Prices", type="o")
ggseasonplot(b1_ts, main="1 Bedroom Listing Sales Prices Seasonality")
b1.decomp = decompose(b1_ts, type = "additive")
plot(b1.decomp)
```
There is an upward trend for the 1 Bedroom time series. Yearly seasonality can be observed, where there are 3 peaks per year, which means the prices may undergo a change every 4 months.


Next, we determine 1 Bedroom Time Series Stationarity:
```{r}
b1_acf = acf2(b1_ts, max.lag = 100)
```
The 1 Bedroom time series is not stationary, but we need to stationarize the data in order to find a model that can simulate the data's behavior.

First order Differencing seems to stationarize the time series.
```{r}
b1_acf_diff = acf2(diff(b1_ts), max.lag = 100)
```
ACF and PACF observations:
1 Bedroom time series seems to be an ARMA(1,1) model since both the ACF and the PACF tail off. Some seasonality can be observed in the plots.

By using EACF we can find an approximately closer (p,q) for the ARMA model seen in the ACF and PACF plots:
```{r}
eacf(diff(b1_ts))
```
From EACF plot we can see the following model possibilities for the lower (p,q) of an ARMA model:
- ARMA(1,3)
- ARMA(1,4)

Given the previously observed seasonality, we will perform EDA and add seasonality to our model.
```{r}
b1_seasonal = acf2(diff(diff(b1_ts), 12), max.lag = 100)
```

After incorporating seasonality for the 1 Bedroom data, it looks like there is a change every 3-4 months and therefore could follow a quarterly seasonality where s = 4. From the ACF and PACF plots we can also observe an ARMA(0,4) model, where ACF cuts off after q = 4, and PACF tails off.

Again, we will look at EACF for possible (P,Q) values for our seasonal model:
```{r}
eacf(diff(diff(b1_ts), 4))
```
From EACF plot we can see the following model possibilities for the capital (P,Q) of a seasonal ARIMA model:
- ARIMA(0,1,4)
- ARIMA(1,1,4)


The model options we got from using ACF and EACF were tested, and after looking at the p-values and modifying the ARIMA (p,d,q) x (P,D,Q) and seasonality to improve the goodness of fit, we found the following model delivers the best AIC, AICc, BIC, and p-values.

1 BEDROOM - ARIMA(1,1,3) x (0,1,3) s = 4
```{r}
sarima(b1_ts, 1,1,3, 0,1,3, 4, no.constant = TRUE, details = FALSE)
```


We still want to apply auto.arima to see if it delivers a better model than ours:
```{r}
auto.arima(b1_ts)
```


Using sarima, we modeled the data using the model provided by auto.arima.
```{r}
# Bedroom 1
sarima(b1_ts, 4,1,2, 0,0,2, 12, no.constant = TRUE, details = FALSE)
```


The model generated by auto.arima did not perform as well, so we modified the seasonality to reflect the one used on our model:
```{r}
# Bedroom 1
sarima(b1_ts, 4,1,2, 0,0,2, 4, no.constant = TRUE, details = FALSE)
```

Our model has the best fit for the 1 Bedroom data, however we will use the 3 models to forecast for the next 12 months so that we can compare the predicted values to the actual values we have for June-September 2018.

Our model:
```{r}
fcast = sarima.for(b1_ts, 12, 1,1,3, 0,1,3, 4, no.constant = TRUE)
```

Auto Arima model with yearly seasonality:
```{r}
fcast_aa = sarima.for(b1_ts, 12, 4,1,2, 0,0,2, 12, no.constant = TRUE)
```

Auto Arima model with quarterly seasonality:
```{r}
fcast_aa4 = sarima.for(b1_ts, 12, 4,1,2, 0,0,2, 4, no.constant = TRUE)
```


Compare forecasted values with actual values from the 1 Bedroom dataset:
```{r}
pred_fcast = fcast$pred[1:4]
pred_fcast_aa = fcast_aa$pred[1:4]
pred_fcast_aa4 = fcast_aa4$pred[1:4]
actual = b1_compare$`77019`
df = data.frame("Actual Price" = actual, "Our Model Forecast" = pred_fcast, "AA Model Forecast" = pred_fcast_aa, "AA Model 2 Forecast" = pred_fcast_aa4)
df
```

```{r}
ts.plot(df, type="o", col= 1:4)
legend("bottomleft", colnames(df), col = 1:4, lty = 1)
```
After comparing the forecasted values from the 3 models against the 1 Bedroom Sales Prices for June-September 2018, the following can be observed:
- The auto arima model with yearly seasonality does not perform as well as we expected from it's lower AIC, AICc, and BIC values than the auto arima quarterly seasonal model.
- The auto arima model with quarterly seasonality forecasted sales price values closest to the actual values.
- Our model's forecasted values were not as close to the actual values as the auto arima quarterly seasonal model, but our model followed the trend the closest.


```{r}
u = fcast$pred[1:4] + (1.96 * fcast$se[1:4])
l = fcast$pred[1:4] - (1.96 * fcast$se[1:4])
df2 = data.frame("Actual Values"=actual, "Fcast Values"=pred_fcast, "Upper"=u, "Lower"=l)
df2
ts.plot(df2, type="o", col= 1:4)
legend("bottomleft", colnames(df2), col = 1:4, lty = 1)
```

As a result we will use our model to forecast since it follows the actual data's trend more closely and the actual data's values still fall within the prediction 95% interval of the forecasted values.

The following are the forecasted values for the next 12 months of 1 Bedroom Listing Sales Prices in the 77019 zip code:
```{r}
fcast$pred
```


Conclusion

We were able to find a model that forecasted the sales prices for 1 Bedroom Listings in the 77019 zip code for 12 months, 4 months of which the forecasted values were close to the actual values provided in the dataset.

The 1 Bedroom Sales Listing Prices started falling in 2006 all the way through to 2013 where they started to rise quickly. Around 2016 the sales prices stopped rising, and instead followed a more seasonal trend. As initially hypothesized, the data was seasonal - quarterly to be more specific - and the listings experience an increase in sales prices during the summer months.


References

Zillow Research:
https://www.zillow.com/research/data/





